Manuscript

### 2022-03-07 analysis/manuscript.Rmd
wflow_publish("analysis/manuscript.Rmd")

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(kableExtra)
library(DT)
library(knitr)
library(ggplot2)
library(kableExtra)
library(cowplot)
library(plotly)
library(plyr)
library(dplyr)
library(ggrepel)
library(viridis)
library(hrbrthemes)
library(ggpubr)
library(broom)
```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(couple_MR_vs_trait_corr)

tar_load(corr_impact_by_PCs)
tar_load(corr_impact_by_traits)
tar_load(corr_impact_by_coords)
tar_load(corr_potential_trait_confounders)
tar_load(traits_corr)
tar_load(PCs_corr)
tar_load(traits_corr2_filled)
tar_load(household_MR_summary_AM)
tar_load(num_tests_by_PCs)
tar_load(num_tests_by_PCs_AM_sig)
tar_load(household_MR_binned_het_sex_spec)
tar_load(household_MR_binned_het_SNPmeta)
tar_load(traits_final)
tar_load(household_MR_AM_FvsM_fig)
tar_load(confounder_traits)
tar_load(proxyMR_prod_comparison_fig_yiyp_adj)
tar_load(proxyMR_comparison_summary_yiyp_adj_SNPmeta)
tar_load(household_MR_summary_AM)
tar_load(household_MR_binned_het_SNPmeta)

@@

@@

Sought to compare the raw correlation amongst couples vs. the standardized MR effects. There are many traits where the `correlation > MR effects` including some traits of interest such as standing height, place of birth, among many others. A plot comparing the two and a summary table of the significant effects is shown below.

@@

interesting_phenotypes <- c("Basal metabolic rate", "Place of birth in UK - north co-ordinate", 
                            "Systolic blood pressure, automated reading", "Standing height", 
                            "Body fat percentage", "Forced vital capacity (FVC), Best measure","Red blood cell (erythrocyte) count")

plot_data <- replace_Neale_ID(couple_MR_vs_trait_corr, traits_corr2_filled, "exposure_ID", "outcome_description") %>% 
  mutate(Phenotype = outcome_description) %>% 
  mutate(Phenotype = case_when(Phenotype == "Basal metabolic rate" ~ "BMR",
                               Phenotype == "Systolic blood pressure, automated reading" ~ "SBP",
                               Phenotype == "Body fat percentage" ~ "BFP",
                               Phenotype == "Forced vital capacity (FVC), Best measure" ~ "FVC",
                               Phenotype == "Red blood cell (erythrocyte) count" ~ "RBC count",
                               Phenotype == "Place of birth in UK - north co-ordinate" ~ "Place of birth, NC",
                               TRUE ~ Phenotype)) %>%
  dplyr::rename(Correlation = couple_r) %>% dplyr::rename(MR_est = IVW_beta) %>% 
  mutate(diff_p_sig = case_when(diff_p < 0.05/num_tests_by_PCs ~ TRUE, TRUE ~ FALSE)) %>%
  mutate(label_phenos = case_when(outcome_description %in% interesting_phenotypes ~ TRUE, TRUE ~ FALSE))

@@

plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = Correlation,
                                                       y = MR_est, 
                                                       label = Phenotype, 
                                                       color = factor(diff_p_sig)),) +
  geom_point(alpha = 1.2) +
  scale_color_manual(values = custom_col[c(1,2)], labels=c("Non-significant","BF-signifncant difference")) +
  #geom_text(show.legend = FALSE) +
  geom_abline(slope=1, intercept=0) +

@@

  theme_minimal() +
  ggplot2::labs(x = paste0("Couple correlation"),
                y = paste("Standardized MR effect")) +
  theme(legend.position="bottom") + theme(legend.title=element_blank()) +
  geom_text_repel(data = subset(plot_data, label_phenos==TRUE),
                  nudge_y = -0.25 - subset(plot_data, label_phenos==TRUE)$MR_est, size = 2,
                  direction = "x")

plot

plot_data.DT <- plot_data %>% dplyr::rename(IVW_beta = MR_est) %>% 
  filter(diff_p < 0.05/num_tests_by_PCs) %>%
  mutate(mr_corr_diff = Correlation - IVW_beta) %>% 
  mutate(correlation_larger = ifelse(Correlation > IVW_beta, TRUE, FALSE)) %>%
  dplyr::select(outcome_description, Correlation, IVW_beta, IVW_pval, diff_p, mr_corr_diff)

@@

## Summary of cases where correlation > MR estimate.

The table below displays the `r dim(plot_data.DT)[1]` traits that have correlation significantly > MR-estimate.

format_round_cols <- colnames(dplyr::select_if(plot_data.DT, is.numeric))

DT::datatable(plot_data.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
```

## Search for potential confounders to explain cases where correlation > MR estimate.

1. Restrict to only traits that have correlation significantly > MR-estimate (`r dim(plot_data.DT)[1]` traits have correlation > MR_estimate).
2. For each of these traits, find the traits which have a causal effect on this trait (MR p-val < 0.05/[number of effective tests]).
3. Calculate the correlation due to confounding as: $C = (\alpha_{z\rightarrow x})^2 * \alpha_{z_i\rightarrow z_p}$.
4. Calculate the ratio of this correlation ($C$) and the correlation of $X$ in partners ($C/r_x$).

A summary of the these results can be seen below:

@@

number_of_potential_confounders <- dim(corr_potential_trait_confounders)[1]/dim(plot_data.DT)[1]

@@

sig_confounders <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
  #mutate(potential_confounders = paste0(exposure_description, collapse = "; ")) %>%
  dplyr::summarize(description = outcome_description[1], n = n(), 
                   potential_confounders = paste0(exposure_description, collapse = "; "), 
                   mean_C_ratio = mean(corr_due_to_confounding_ratio)) %>%
  dplyr::rename(trait_ID = outcome_ID)

confounder_summary <- rbind(non_sig_confounders, sig_confounders)

format_round_cols <- colnames(dplyr::select_if(confounder_summary, is.numeric))

DT::datatable(confounder_summary, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)

@@

AM_result_sig.DT <- AM_result_sig %>% dplyr::select(exposure_description, IVW_beta, IVW_pval) %>% unique()
DT::datatable(AM_result_sig.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval'), digits=3)

@@

- For each AM MR ($X_{p} \sim X_i$), analyses were run in each sex separately. 
- Of the `r dim(AM_sig_traits)[1]` significant results shown above, we tested to see if there was any difference in AM MR estimates between sexes. 
- A similar adjustment for multiple hypothesis testing was applied as above, resulting in `r num_tests_by_PCs_AM_sig` number of effective tests. 
- After adjusting for multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig``), `r dim(AM_sex_spec_result_sig)[1]` traits showed significant differences amongst sexes (this changed from before).
- The table below shows the *nominally significant* results at `p < 0.05`.

@@

AM_sig_sex_het_nom <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05)
AM_sig_sex_het_nom <- replace_Neale_ID(AM_sig_sex_het_nom, traits_corr2_filled,"exposure_ID", "exposure_description")
AM_result_sig_nom.DT <- AM_sig_sex_het_nom %>% dplyr::select(exposure_description, exposure_sex, IVW_beta, IVW_pval, sex_het_pval) %>% unique()
DT::datatable(AM_result_sig_nom.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'sex_het_pval'), digits=3)

## Binned results (exploring effect of age and time-together).

@@

sex_het_wt_sig_age_traits <- replace_Neale_ID(sex_het_wt_sig_age, traits_corr2_filled, "exposure_ID", "outcome_description") %>% pull(outcome_description) %>% unique()
```

@@

Given that we have no significant results, how should we report these analyses?

A few potential figures we could include for a given trait are shown below. 
```{r binned_figures, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = TRUE}

tar_load(exposure_info)
for(i in 1:length(exposure_info)){
  temp <- exposure_info[[i]]
  if(temp[which(temp$Value=="trait_ID"), "Info"] == "130_irnt") {break}
}

figs <- tar_read(household_MR_binned_AM_figs, branches = i)[[1]]

figs[[1]]
figs[[2]]
figs[[3]]
figs[[4]]
figs[[5]]
figs[[6]]
figs[[7]]
figs[[8]]
figs[[9]]

>>>>

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(kableExtra)
library(DT)
library(knitr)
library(ggplot2)
library(kableExtra)
library(cowplot)
library(plotly)
library(plyr)
library(dplyr)
library(ggrepel)
library(viridis)
library(hrbrthemes)
library(ggpubr)
library(broom)
library(data.table)
library(xlsx)
```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

tar_load(couple_MR_vs_trait_corr)
tar_load(corr_impact_by_PCs)
tar_load(corr_impact_by_traits)
tar_load(corr_impact_by_coords)
tar_load(corr_potential_trait_confounders)
tar_load(traits_corr)
tar_load(PCs_corr)
tar_load(traits_corr2_filled)
tar_load(household_MR_summary_AM)
tar_load(num_tests_by_PCs)
tar_load(num_tests_by_PCs_AM_sig)
tar_load(household_MR_binned_het_sex_spec)
tar_load(household_MR_binned_het_SNPmeta)
tar_load(traits_final)
tar_load(household_MR_AM_FvsM_fig)
tar_load(confounder_traits)
tar_load(proxyMR_prod_comparison_fig_yiyp_adj)
tar_load(proxyMR_comparison_summary_yiyp_adj_SNPmeta)
tar_load(household_MR_summary_AM)
tar_load(household_MR_binned_het_SNPmeta)
tar_load(potential_trait_confounders_MVMR)
tar_load(corr_mat_traits)

@@

# Pearson vs. Spearman correlation.

We wanted to ensure that INQT was not significantly impacting the correaltions of each trait, so we also calculated a non-parametric based correaltion (Spearman) and found consistent correlations.

```{r pearson_vs_spearman, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE, warning = FALSE}
tar_load(Neale_SGG_dir_filt)
tar_load(traits_corr_spearman)
tar_load(traits_corr)

Neale_SGG_dir_filt_unique <- Neale_SGG_dir_filt[,c("phenotype", "phenotype_sub", "sgg_phesant_name", "description", "SGG_available", "variable_type")] %>% unique()

pearson_vs_spearman_dat <- inner_join(traits_corr, unique(Neale_SGG_dir_filt_unique[,c("sgg_phesant_name", "variable_type")]), by = c("ID"="sgg_phesant_name")) %>% dplyr::rename(r2_pearson = r2) %>%
  inner_join(traits_corr_spearman[,c("ID", "r2")]) %>% dplyr::rename(r2_spearman = r2) %>% 
  as_tibble() %>% type_convert() %>%
  mutate(r_pearson = sqrt(r2_pearson)) %>% 
  mutate(r_spearman = sqrt(r2_spearman))

pearson_vs_spearman_plot <- ggplot2::ggplot(data = pearson_vs_spearman_dat, 
                                            ggplot2::aes(x = r_pearson,
                                                         y = r_spearman, 
                                                         color = factor(variable_type))) + 
  geom_point() +
  theme_minimal() +
  ggplot2::labs(x = paste0("Pearson correlation"),
                y = paste("Spearman correlation")) + 
  scale_color_manual(name = "Variable type", values = custom_col[c(1,2,3)], labels=c("Binary","Continuous (IRNT)", "Ordinal"))

pearson_vs_spearman_plot
```

@@

Sought to compare the raw correlation amongst couples vs. the standardized MR effects. There are many traits where the `correlation > MR effects` including some traits of interest such as standing height, place of birth, among many others.

For example, if we take a trait such as place of birth (NC), this is very highly correlated among couples. However, the within-couple MR-estimate is smaller than the observed phenotypic correaltion. This is the case for many such traits.

```{r pheno_corr_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE, warning = FALSE}

### phenotypic correlation plot for 129_irnt (place of birth)

tar_load(path_pheno_data)
tar_load(hh_pairs_filter)
Neale_pheno_ID <- "129_irnt" ## this is the Neale_id, used to be pheno_description
phes_ID <- gsub("_irnt", "", Neale_pheno_ID)

male_pheno_file <- path_pheno_data[which(endsWith(path_pheno_data, paste0("/phesant/", Neale_pheno_ID, "_male.txt")))]
female_pheno_file <- path_pheno_data[which(endsWith(path_pheno_data, paste0("/phesant/", Neale_pheno_ID, "_female.txt")))]
male_pheno_data <- fread(male_pheno_file, header=TRUE, select=c("IID","sex","age", phes_ID))
female_pheno_data <- fread(female_pheno_file, header=TRUE, select=c("IID","sex","age", phes_ID))

joint_data <- rbind(male_pheno_data, female_pheno_data)

pairs_filter_copy <- hh_pairs_filter
    pairs_filter_copy <- pairs_filter_copy %>%
      mutate(HOUSEHOLD_MEMBER1_M = case_when(HOUSEHOLD_MEMBER1_sex == 1 ~ HOUSEHOLD_MEMBER1,
                                             HOUSEHOLD_MEMBER2_sex == 1 ~ HOUSEHOLD_MEMBER2)) %>%

      mutate(HOUSEHOLD_MEMBER2_F = case_when(HOUSEHOLD_MEMBER1_sex == 0 ~ HOUSEHOLD_MEMBER1,
                                             HOUSEHOLD_MEMBER2_sex == 0 ~ HOUSEHOLD_MEMBER2)) %>%

      dplyr::select(-HOUSEHOLD_MEMBER1, -HOUSEHOLD_MEMBER2) %>%
      dplyr::rename(HOUSEHOLD_MEMBER1 = HOUSEHOLD_MEMBER1_M) %>%
      dplyr::rename(HOUSEHOLD_MEMBER2 = HOUSEHOLD_MEMBER2_F)

pairs_filter_copy$trait1 <- joint_data[[phes_ID]][match(pairs_filter_copy[["HOUSEHOLD_MEMBER1"]], joint_data$IID)]
pairs_filter_copy$trait2 <- joint_data[[phes_ID]][match(pairs_filter_copy[["HOUSEHOLD_MEMBER2"]], joint_data$IID)]

pheno_corr_dat <- pairs_filter_copy

pheno_corr_plot <- ggplot2::ggplot(data = pheno_corr_dat, ggplot2::aes(x = trait1,
                                                       y = trait2)) +
  geom_point(alpha = 1.2, color = "gray") +
  #geom_text(show.legend = FALSE) +
  #geom_abline(slope=1, intercept=0) +

  theme_minimal() +

  #geom_text(aes(x=1,y=2,label=""),show.legend=F) +
  ggplot2::labs(x = paste0("Male place of birth (NC)"),
                y = paste("Female place of birth (NC)")) + 
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()

pheno_corr_plot

A plot comparing the two effects, and a summary table of the significant effects is shown below.

@@

interesting_phenotypes <- c("Basal metabolic rate", "Place of birth in UK - north co-ordinate", 
                            "Systolic blood pressure, automated reading", "Standing height", 
                            "Weight", "Forced vital capacity (FVC), Best measure",
                            "Basophill count", "Pulse rate, automated reading")

corr_vs_MR_effects_dat <- replace_Neale_ID(couple_MR_vs_trait_corr, traits_corr2_filled, "exposure_ID", "outcome_description") %>% 
  mutate(Phenotype = outcome_description) %>% 
  mutate(Phenotype = case_when(Phenotype == "Basal metabolic rate" ~ "BMR",
                               Phenotype == "Systolic blood pressure, automated reading" ~ "SBP",
                               Phenotype == "Pulse rate, automated reading" ~ "Pulse rate",
                               Phenotype == "Body fat percentage" ~ "BFP",
                               Phenotype == "Standing height" ~ "Height",
                               Phenotype == "Reticulocyte percentage" ~ "Reticulocyte %",

                               #Phenotype == "Weight" & outcome_ID == "21002_irnt" ~ "Weight",
                               Phenotype == "Forced vital capacity (FVC), Best measure" ~ "FVC",
                               Phenotype == "Red blood cell (erythrocyte) count" ~ "RBC count",
                               Phenotype == "Place of birth in UK - north co-ordinate" ~ "Place of birth, NC",
                               TRUE ~ Phenotype)) %>%

  dplyr::rename(Correlation = couple_r) %>% dplyr::rename(MR_est = IVW_beta) %>% 
  mutate(diff_p_sig = case_when(diff_p < 0.05/num_tests_by_PCs ~ TRUE, TRUE ~ FALSE)) %>%
  mutate(outcome_description2 = case_when(outcome_description=="Weight" & outcome_ID == "23098_irnt" ~ "Weight2",
                                          TRUE ~ outcome_description)) %>%
  mutate(label_phenos = case_when(outcome_description2 %in% interesting_phenotypes ~ TRUE, TRUE ~ FALSE))

@@

corr_vs_MR_effects_plot <- ggplot2::ggplot(data = corr_vs_MR_effects_dat, ggplot2::aes(x = Correlation,
                                                       y = MR_est, 
                                                       label = Phenotype, 
                                                       color = factor(diff_p_sig)),) +
  geom_point(alpha = 1.2) +
  scale_color_manual(values = custom_col[c(1,2)], labels=c("Non-significant","BF-signifncant difference")) +
  #geom_text(show.legend = FALSE) +
  geom_abline(slope=1, intercept=0) +

@@

  theme_minimal() +
  theme(legend.text=element_text(size=11)) +
  #geom_text(aes(x=1,y=2,label=""),show.legend=F) +
  ggplot2::labs(x = paste0("Couple correlation"),
                y = paste("Standardized MR effect")) +
  theme(legend.position="bottom") + theme(legend.title=element_blank()) +
  geom_text_repel(data = subset(corr_vs_MR_effects_dat, label_phenos==TRUE),
                  nudge_y = -0.25 - subset(corr_vs_MR_effects_dat, label_phenos==TRUE)$MR_est, size = 3.25,
                  direction = "x", show.legend = FALSE )

corr_vs_MR_effects_plot

corr_vs_MR_effects_sig_dat <- corr_vs_MR_effects_dat %>% dplyr::rename(IVW_beta = MR_est) %>% 
  filter(diff_p < 0.05/num_tests_by_PCs) %>%
  mutate(mr_corr_diff = Correlation - IVW_beta) %>% 
  mutate(correlation_larger = ifelse(Correlation > IVW_beta, TRUE, FALSE)) %>% 
  arrange(diff_p) %>% dplyr::select(-exposure_ID, -Phenotype, -outcome_description2, -label_phenos, -mr_corr_diff, -diff_p_sig) %>% 
  dplyr::rename(trait_description = outcome_description) %>%
  dplyr::rename(trait_ID = outcome_ID) %>%
  dplyr::rename(couple_r = Correlation) %>%
  dplyr::select(trait_description, trait_ID, everything())

## prune for correlations < 0.8

corr_vs_MR_effects_sig_dat.DT <- corr_vs_MR_effects_sig_dat %>% 
  dplyr::select(trait_description, trait_ID, couple_r, IVW_beta, IVW_pval, diff_p, correlation_larger)

#prune_pheno_table(corr_vs_MR_effects_sig_dat, "trait_ID", corr_mat_traits, corr_trait_threshold)

# Could prune the results, but only one trait is removed so seems pointless:

# Result are pruned to only show traits with correlation < `r corr_trait_threshold`. (The number of traits before pruning was `r dim(corr_vs_MR_effects_sig_dat)[1]` vs `r dim(corr_vs_MR_effects_sig_dat_pruned.DT)[1]` after pruning).
```

## Summary of cases where correlation is different from the MR estimate.

The table below displays the `r dim(corr_vs_MR_effects_sig_dat.DT)[1]` traits that have correlation significantly different from the MR-estimate. Of the significant difference `r table(corr_vs_MR_effects_sig_dat$correlation_larger)[[1]]` have MR-estimates larger than than correlation, suggesting the presence of a negative confounder.

format_round_cols <- colnames(dplyr::select_if(corr_vs_MR_effects_sig_dat.DT, is.numeric))

DT::datatable(corr_vs_MR_effects_sig_dat.DT %>% dplyr::select(-trait_ID), width = "100%") %>% formatSignif(columns=format_round_cols, digits = 3)

@@

## Search for potential confounders to explain cases where correlation differs from MR estimate.

1. Restrict to only traits that have correlation significantly different from MR-estimate (`r dim(corr_vs_MR_effects_sig_dat)[1]` traits have correlation different from MR_estimate).
2. For each of these traits, find the traits which have a causal effect on this trait (MR p-val < 0.05/[number of effective tests]).
3. Calculate the correlation due to confounding as: $C = (\alpha_{z\rightarrow x})^2 * \alpha_{z_i\rightarrow z_p}$.
4. Calculate the ratio of this correlation ($C$) and the correlation of $X$ in partners ($C/r_x$). 
5. Remove any `potential_counders` with correlation with `trait_ID` > `r corr_trait_threshold`.
6. For each `trait_ID` prune remaining `potential_counders` so that all have correlations <  `r corr_trait_threshold`, prioritized by their `corr_due_to_confounding_ratio`.

A summary of the top 5 confounders for each `trait_ID` can be seen below:

```{r simple_C_model, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/simple_confounder_model.png", error = FALSE)

@@

number_of_potential_confounders <- dim(corr_potential_trait_confounders)[1]/dim(corr_vs_MR_effects_sig_dat)[1]

@@

sig_confounders <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
  arrange(-corr_due_to_confounding_ratio) %>%
  mutate(total_potential_confounders = n()) %>%
  dplyr::summarize(description = outcome_description[1],
                   total_potential_confounders = total_potential_confounders[1],
                   potential_confounders = paste0(exposure_description, collapse = "; "), 
                   max_C = max(corr_due_to_confounding), 
                   min_C = min(corr_due_to_confounding)) %>%
  dplyr::rename(trait_ID = outcome_ID) %>%
  dplyr::rename(trait_description = description) %>% 
  dplyr::select(trait_description, trait_ID, everything())

sig_confounders_top5 <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
  arrange(-corr_due_to_confounding_ratio) %>%
  mutate(total_potential_confounders = n()) %>%
  slice_head(n = 5) %>%
  #mutate(potential_confounders = paste0(exposure_description, collapse = "; ")) %>%
  dplyr::summarize(trait_description = outcome_description[1],
                   total_potential_confounders = total_potential_confounders[1],
                   top5_potential_confounders = paste0(exposure_description, collapse = "; "), 
                   max_C = max(corr_due_to_confounding), 
                   min_C = min(corr_due_to_confounding)) %>%
  dplyr::rename(trait_ID = outcome_ID) %>%
  left_join(corr_vs_MR_effects_sig_dat.DT, by = c("trait_description", "trait_ID")) %>%
  dplyr::select(trait_description, trait_ID, couple_r, IVW_beta, IVW_pval,everything())

format_round_cols <- colnames(dplyr::select_if(sig_confounders_top5, is.numeric))

DT::datatable(sig_confounders_top5, width = "100%") %>% formatSignif(columns=format_round_cols, digits = 3)

We compared the difference in phenotypic estimate and MR-estimates to the maximum C for each trait, as shown in the figure below:

```{r max_C_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

max_C_data <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
  arrange(-corr_due_to_confounding_ratio) %>%
  mutate(total_potential_confounders = n()) %>%
  slice_head(n = 1) %>%
  left_join(couple_MR_vs_trait_corr %>% 
              dplyr::rename(outcome_ID_IVW_beta = IVW_beta) %>%
              dplyr::rename(outcome_ID_IVW_se = IVW_se) %>%
              dplyr::rename(outcome_ID_couple_r = couple_r) %>% 
              dplyr::rename(outcome_ID_couple_r_se = couple_r_se) %>% 
              dplyr::select(outcome_ID, outcome_ID_IVW_beta, outcome_ID_IVW_se, outcome_ID_couple_r, outcome_ID_couple_r_se)) %>%
  mutate(same_trait_corr_minus_MR_est = abs(outcome_ID_couple_r)- abs(outcome_ID_IVW_beta)) %>%

  dplyr::rename(Z_beta = IVW_beta) %>%
  dplyr::rename(Z_se = IVW_se) %>%
  dplyr::rename(Z_pval = IVW_pval) %>%
  mutate(Z_sq_beta = Z_beta^2) %>%
  mutate(Z_sq_se = sqrt(4*Z_beta^2*Z_se^2+2*Z_se^4)) %>%
  dplyr::rename(Z_AM_beta = exposure_ID_AM_IVW_beta) %>%
  dplyr::rename(Z_AM_se = exposure_ID_AM_IVW_se) %>%
  mutate(Z_prod = Z_sq_beta*Z_AM_beta) %>%
  mutate(Z_prod_var = variance_of_product(Z_sq_beta, Z_sq_se, Z_AM_beta, Z_AM_se)) %>%
  mutate(Z_prod_se = sqrt(Z_prod_var)) %>%
  mutate(same_trait_corr_minus_MR_est_se = variance_of_sum(outcome_ID_IVW_se, outcome_ID_couple_r_se))

max_C_plot <- ggplot2::ggplot(data = max_C_data, ggplot2::aes(x = corr_due_to_confounding,
                                                       y = same_trait_corr_minus_MR_est)) +
  geom_point(alpha = 1.2) +
  #scale_color_manual(values = custom_col[c(1,2)], labels=c("Non-significant","BF-signifncant difference")) +
  #geom_text(show.legend = FALSE) +
  geom_abline(slope=1, intercept=0) +

  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = same_trait_corr_minus_MR_est - same_trait_corr_minus_MR_est_se, ymax = same_trait_corr_minus_MR_est + same_trait_corr_minus_MR_est_se),
    colour = "grey",
    width = 0
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(xmin = corr_due_to_confounding - Z_prod_se, xmax = corr_due_to_confounding + Z_prod_se),
    colour = "grey",
    width = 0
  ) +

  theme_minimal() +
  ggplot2::labs(x = paste0("max C"),
                y = paste("abs(correlation) - abs(MR-estimate)")) +
  theme(legend.title=element_blank()) +
  #geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()

max_C_plot

Finally, we computed a MVMR to sum up the effect of all potential confounders as shown in the DAG below and compared these C-statistics to the difference between phenotypic correlation and MR-estimates.

```{r MVMR_C_model, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/MVMR_confounder_model.png", error = FALSE)
```

```{r MVMR_C_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

discrepentant_traits <- unique(corr_potential_trait_confounders$outcome_ID)

output <- numeric()
for(i in 1:length(potential_trait_confounders_MVMR)){

  outcome_ID <- discrepentant_traits[i]

  corr_potential_trait_confounders_i <- corr_potential_trait_confounders %>% filter(outcome_ID == !!outcome_ID)

  MVMR_result_org <- potential_trait_confounders_MVMR[[i]]

  MVMR_result_org <- MVMR_result_org %>% mutate(beta_2_var = 4*beta^2*se^2+2*se^4) %>%
    mutate(beta_2_se = sqrt(beta_2_var)) %>% 
    left_join(corr_potential_trait_confounders_i %>% dplyr::select(exposure_ID, exposure_ID_AM_IVW_beta, exposure_ID_AM_IVW_se), by = c("exposure" = "exposure_ID")) %>%
    dplyr::rename(MVMR_Z_beta = beta) %>%
    dplyr::rename(MVMR_Z_se = se) %>%
    dplyr::rename(MVMR_Z_pval = pval) %>%
    mutate(MVMR_Z_sq_beta = MVMR_Z_beta^2) %>%
    dplyr::rename(MVMR_Z_sq_se = beta_2_se) %>% 
    dplyr::rename(Z_AM_beta = exposure_ID_AM_IVW_beta) %>%
    dplyr::rename(Z_AM_se = exposure_ID_AM_IVW_se) %>%
    mutate(Z_prod = MVMR_Z_sq_beta*Z_AM_beta) %>%
    mutate(Z_prod_var = variance_of_product(MVMR_Z_sq_beta, MVMR_Z_sq_se, Z_AM_beta, Z_AM_se))

  MVMR_result_nom <- MVMR_result_org %>% filter(MVMR_Z_pval < 0.05)
  MVMR_result_BF <- MVMR_result_org %>% filter(MVMR_Z_pval < 0.05/dim(MVMR_result_org)[1])

  num_confounders_nom <- dim(MVMR_result_org)[1]
  num_confounders_BF <- dim(MVMR_result_BF)[1]

  row <- cbind(outcome_ID)
  for(filter_type in c("nom", "BF")){

    if(filter_type == "nom"){
      t <- MVMR_result_org %>% filter(MVMR_Z_pval < 0.05)
    }
    if(filter_type == "BF"){
      t <- MVMR_result_org %>% filter(MVMR_Z_pval < 0.05/dim(MVMR_result_org)[1])
    }

    num_confounders <- dim(t)[1]
    C_sum <- sum(t$Z_prod)
    C_sum_var <- sum(t$Z_prod_var)
    C_sum_se <- sqrt(C_sum_var)

    temp_row <- cbind(num_confounders, C_sum, C_sum_var, C_sum_se)
    colnames(temp_row) <- paste0(c("num_confounders", "C_sum", "C_sum_var", "C_sum_se"), "_", filter_type)
    row <- cbind(row, temp_row)

  output <- rbind(output, row)

output <- output %>% as_tibble() %>% type_convert() %>% mutate(C_sum_pval = 2*pnorm(-abs(C_sum_nom/C_sum_se_nom)))

MVMR_C_data <- max_C_data %>% left_join(output)

MVMR_C_plot <- ggplot2::ggplot(data = MVMR_C_data, ggplot2::aes(x = C_sum_nom ,
                                                       y =  same_trait_corr_minus_MR_est)) +
  geom_point(alpha = 1.2) +
  #scale_color_manual(values = custom_col[c(1,2)], labels=c("Non-significant","BF-signifncant difference")) +
  #geom_text(show.legend = FALSE) +
  #geom_abline(slope=1, intercept=0) +

  # ggplot2::geom_errorbar(
  #   ggplot2::aes(ymin = MR_est - IVW_se, ymax = MR_est + IVW_se),
  #   colour = "grey",
  #   width = 0
  # ) +
  # ggplot2::geom_errorbar(
  #   ggplot2::aes(xmin = Correlation - couple_r_se, xmax = Correlation + couple_r_se),
  #   colour = "grey",
  #   width = 0
  # ) +

  theme_minimal() +
  ggplot2::labs(x = paste0("MVMR C-sum (p < 0.05)"),
                y = paste("abs(correlation) - abs(MR-estimate)")) +
  theme(legend.title=element_blank()) +
  geom_abline(slope=1, intercept=0)

MVMR_C_plot

@@

AM_result_sig.DT <- AM_result_sig %>% dplyr::select(exposure_description, exposure_ID, IVW_beta, IVW_se, IVW_pval) %>% unique() %>% 
  dplyr::rename(trait_description = exposure_description) %>%
  dplyr::rename(trait_ID = exposure_ID)

DT::datatable(AM_result_sig.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_se', 'IVW_pval'), digits=3)

AM_sig_sex_het_nom <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05)
AM_sig_sex_het_nom <- replace_Neale_ID(AM_sig_sex_het_nom, traits_corr2_filled,"exposure_ID", "exposure_description") %>%
  dplyr::rename(trait_description = exposure_description) %>%
  dplyr::rename(trait_ID = exposure_ID) %>%
  dplyr::select(trait_description, trait_ID, everything())

AM_result_sig_nom.DT <- AM_sig_sex_het_nom %>% dplyr::select(trait_description, trait_ID, exposure_sex, IVW_beta, IVW_se, IVW_pval, sex_het_pval) %>% unique()
```

@@

- For each AM MR ($X_{p} \sim X_i$), analyses were run in each sex separately. 
- Of the `r dim(AM_sig_traits)[1]` significant results shown above, we tested to see if there was any difference in AM MR estimates between sexes. 
- A similar adjustment for multiple hypothesis testing was applied as above, resulting in `r num_tests_by_PCs_AM_sig` number of effective tests. 
- After adjusting for multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig``), `r dim(AM_sex_spec_result_sig)[1]` traits showed significant differences amongst sexes.
- The table below shows the *nominally significant* results at `p < 0.05` (of which there are `r dim(AM_result_sig_nom.DT)[1]/2`).

@@

DT::datatable(AM_result_sig_nom.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'sex_het_pval'), digits=3)

nominally_sig_sex_differences <- AM_result_sig_nom.DT %>% pivot_wider(names_from = exposure_sex, values_from = c(IVW_beta, IVW_se, IVW_pval)) 
# write.table(nominally_sig_sex_differences, "docs/manuscript_tables/nominally_sig_sex_differences.txt", row.names = F, sep=",", quote = F)
```

## MR binned results (exploring effect of age and time-together).

@@

@@

The significant traits correspond to the following traits:

```{r sig_sex_diff, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = TRUE}

sex_het_wt_sig_age_traits <- replace_Neale_ID(sex_het_wt_sig_age, traits_corr2_filled, "exposure_ID", "outcome_description") %>% pull(outcome_description) %>% unique()

sex_het_sig_age_sub <- sex_het_sig_age %>% dplyr::select(exposure_ID, exposure_sex, grouping_var, bin_slope_beta, bin_slope_se, bin_slope_pval, bin_slope_sex_het_pval)
sex_het_sig_tt_sub <- sex_het_sig_tt %>% dplyr::select(exposure_ID, exposure_sex, grouping_var, bin_slope_beta, bin_slope_se, bin_slope_pval, bin_slope_sex_het_pval)

sig_sex_diff_dat <- replace_Neale_ID(rbind(sex_het_sig_age_sub, sex_het_sig_tt_sub), traits_corr2_filled, "exposure_ID", "exposure_description")

DT::datatable(sig_sex_diff_dat, width = "100%") %>% formatSignif(columns=c('bin_slope_beta', 'bin_slope_se', 'bin_slope_pval', 'bin_slope_sex_het_pval'), digits=3)

A few potential figures we could include for a given trait are shown below.
